---
title: "W1: 1st meeting & learning the arcane language"
output: html_notebook
---

*First, let us introduce ourselves and what we hope to achieve.  *

**Goals of the group:**  
  
-learn publication-quality statistical analysis and visualization  
-friendly and open environment to increase fellowship  
-locally relevant  
-accessible for beginners & experts  
-teaching/service experience for the CV if you want to lead session(s)  
-others?  
  
  
**Components of sessions:**  
  
-expert-led lesson  
-unstructured period for work and collaborative help, "how do I...?"  
  
  
**Lesson topics:**  
  
*Week 2*  
-best practices for excel files, data management  
-data visualization with ggplot()  
--mean and SEM counts on different dates by one or more factors  
--summaries with bar charts or boxplots  
--subsetting parts of datafiles  
--others?  

*Week 3*  
-Basic statistics, the "dark craft"  
--ANOVA, ANCOVA, linear model, t-test, tukey test, others?  

*Future* 
  
-High priority  
--visualization: creating and saving publication-quality figures with ggplot()  
--dark arts of statistics: generalized linear models with glmmTMB()  
  
-Less priority  
--github for version control, data sharing, and reproducibility  
--if/then, while loops, and for loops  
  
-lowest priorty  
--postmodern statistics: bayesian, PCA, structural equation models  
--databases like SQL  
--map stuff like GIS or similar functions  
  
  
  
Let us begin. This code and associated files can be found in: https://github.com/robertorpet/raindrop

```{r Chunk 1 - using the notebook and the console}
#using R like a calculator
1 + 1
1+1
1                         +1

#storing things in variables
rob_variable <- 1 + 1 #notice use of "<-"
rob_variable
rob_variable <- rob_variable + 1
rob_othervariable = 2 + 2 #you can use "=" but that's not preferred
rob_othervariable
print("this is rob's variable")
rob_variable
rob_variable == rob_othervariable # "==" is to check if it is equal
rob_variable = rob_othervariable 
rob_variable

#operators for logic. These can be useful for subsetting data
1 > 2
1 == 2
1 < 2
1 != 2
# try 1 = 2 or 1 <- 2. you can't assign values to numerals, they are reserved.

#multiple numbers and basic statistics summary
#things that have "()" after them are "functions". Inside the () hit TAB to see what arguments it likes
x <- seq(from = 0, to = 5500, by = 5) # inside functions people tend to use "=" instead of "<-", but the latter works
x <- seq(0,5500, by = 5) # inside functions people tend to use "=" instead of "<-", but the latter works
y <- seq(0,5500, by <- 5)
mean(x)
mean(c(1,2,3,4,5)) # "c()" is for concatenate. mean(), sum(), sd(), length() are all functions that R "knows"
sum(x)
sd(x)
length(x)
SEM <- mean(x)/sqrt(length(x))
SEM
plot(x,y)

#visualize a normal distribution
y <- dnorm(x = x, mean = 2000, sd = 500)
plot(x,y)

#checking class
class(1)
class("a")
b <- "a"
class(b)
class(x)

#check the levels or "unique" values of a factor
rob_listofletters <- c("a", "a", "f", "w", "a", "A", "sequins", "nileseldredge")
class(rob_listofletters)
unique(rob_listofletters)
levels(rob_listofletters) #"levels" doesn't work unless the class is "factor" instead of "character"
rob_listofletters <- as.factor(c("a", "a", "f", "w", "a", "A", "sequins", "nileseldredge"))
unique(rob_listofletters)
levels(rob_listofletters)
my_special_list <- levels(rob_listofletters)
class(my_special_list)
```

```{r Chunk 2 - working with data}
#R comes with some sample datasets you can see by typing data()
data()
data_iris <- iris #iris is one of the built-in datasets
head(data_iris)
names(data_iris)
class(data_iris$Species)
unique(data_iris$Species)
levels(data_iris$Species)
mean(data_iris$Sepal.Length)

#summarize using aggregate(). We are using functions within functions.
data_iris_sepal <- do.call(data.frame, #"do.call(data.frame," is optional but helps put in good format for downstream processing
  aggregate(Sepal.Length~Species,
    data = data_iris,
    function(x)
    c(mean = mean(x),
    count = length(x),
    SEM = sd(x)/sqrt(length(x)))))
data_iris_summary <- do.call(data.frame,
  aggregate(cbind(Sepal.Length,Sepal.Width)~Species, #we  make use of cbind() within aggregate() to summarize multiple "y"s
    data = data_iris,
    function(x)
    c(mean = mean(x),
    count = length(x),
    SEM = sd(x)/sqrt(length(x)))))

#for demo purposes I am going to add a new factor to the dataset to show how to aggregate by multiple things
newcolumn <- rep(c(letters[1:3]),times=50) #this makes a new column with a bunch of letters
data_iris_new <- cbind(data_iris,newcolumn) #use "cbind()" to combine the columns
data_iris_summary_new <- do.call(data.frame,
  aggregate(cbind(Sepal.Length,Sepal.Width)~Species+newcolumn, #on the right side of the "~", you can list multiple things to summarize by using "+"
    data = data_iris,
    function(x)
    c(mean = mean(x),
    count = length(x),
    SEM = sd(x)/sqrt(length(x)))))
```

```{r Chunk 3 - loading your data}
#loading with read.csv()
data <- read.csv("data_week1.csv") #notice the first cell has some weird characters, this the file encoding for csvs and we don't want to read it
data <- read.csv("data_week1.csv", fileEncoding = 'UTF-8-BOM')

#installing and loading packages
library(readxl)

#loading with read_excel()
data <- read_excel("data_week1.xlsx", skip = 9,
guess_max = 10000, #guessmax prevents columns with many initial banks being imported as a logical FALSE for not NAs)
na = " ") # you can put a period in the quotes if that is how you code blanks.

#You can load data from googledocs
#library(googlesheets4)
#Data_scouting <- read_sheet("put the URL here", skip = 0)
```
