---
title: "W4"
output: html_notebook
---

Generalized linear mixed models
Theme: don't use these tools unless you need to.
Items: (1) finish glmm for now. (2) Data sharing.

REVIEW
Reasons to use generalized linear model:
1. binomial ("binary") data (Warton 2011)
2. can't meet assumptions of normal linear model, so you can try other distributions
4. output is related to the world
5. some  fancy options like zero-inflated distributions
6. your advisor or teacher tells you to
7. they make some statisticians happy
8. to feel smart, impress colleagues, employers

Reasons not to use:
1. log-transormation is just as good for most applied statistics (Ives 2015 vs. O'Hara and Kotze 2010)
2. generalized model has higher type I error (false positive) vs. log transform with small sample sizes (Ives 2015)
3. you can find literature for any approach

Summary: you can find support (O'Hara and Kotze 2010) or dissent (Ives 2015) for glmms vs. transformation.


NEW ITEMS
Reasons to used mixed effects model
1. random effects for blocks
2. random effects to control for repeated measurements (prevent pseudoreplication)
3. your advisor tells you to

Reasons not to use:
1. harder to understand and specify model correctly
2. simpler alternatives: you can use a fixed effect for block. You can summarize data for repeated measurements



dealing with pseudoreplication by summarizing (no random effects)
```{r}
#LOAD PACKAGES
library(readxl)
library(ggplot2)

#LOAD DATA
Data <- read_excel("data_week6_PPtapspractice.xlsx")
Data$Site <- as.factor(Data$Site)#convert Site to categorical variable instead of number

#EXPLANATION
#What you should not do: run a model before graphing and thinking
model_anova <- aov(Adults ~ Trt, data = Data)
model_anova <- aov(Adults ~ Trt + Site, data = Data)

summary(model_anova)

#data points are not independent, we only have 6 sites
ggplot(data = Data, aes(x = Trt, y = Adults, color = Site)) +
  geom_point() +
  facet_wrap(~Site)

ggplot(data = Data, aes(x = Trt, y = Adults, color = Site)) +
  geom_point()

#the correct way to look at it. N = 6 paired sites
ggplot(data = Data, aes(x = Trt, y = Adults, color = Site)) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line", aes(group = Site))

#a simple way to deal with this. summarize the data to one value per experimental unit
#aggregate fucntion aggregate()
Data_summary <- do.call(data.frame, 
                        aggregate(Adults~Trt+Site, 
                                  data = Data, 
                                  function(x) 
                                    c(mean = mean(x),
                                      sd = sd(x), 
                                      count = length(x))))

ggplot(data = Data_summary, aes(x = Trt, y = Adults.mean, color = Site)) +
  geom_point() +
  facet_wrap(~Site)

model_anova <- aov(Adults.mean ~ Trt + Site, data = Data_summary) #treating site like a block like in previous example
summary(model_anova)
plot(model_anova, 1)
plot(model_anova, 2)

```


dealing with psuedoreplication by using random effects
```{r}

#LOAD PACKAGES
library(glmmTMB)
library(DHARMa)

#USING RANDOM EFFECTS
#wrong way: ignoring everything. Same wrong approach as above but with glmmTMB()
model <- glmmTMB(formula = Adults ~ Trt            , data = Data) #P is similar to aov example. Sample size is inappropriate
summary(model)
plot(simulateResiduals(model))

#including random effect for repeated measures
#first, let's make an ID for each plot called "subject"
Data$subject <- paste(Data$Site,Data$Trt)
#now add random intercept for subjects
model <- glmmTMB(formula = Adults ~ Trt + (1|subject), data = Data) #we are saying, look there are going to be multiple measurements per subject!
summary(model) #not significant when pseudoreplication is removed
plot(simulateResiduals(model))

#but, we also want to consider that measurements are related by block (i.e., "Site" in this data set)
model <- glmmTMB(formula = Adults ~ Trt + (1|subject) + (1|Site), data = Data)
summary(model) # P is similar to the simple summarized method
plot(simulateResiduals(model)) #let's ignore this for now

#alternative way with "nesting" removes need for subject label
model <- glmmTMB(formula = Adults ~ Trt + (1|Site/Trt), data = Data)
summary(model)
plot(simulateResiduals(model))

#can we satisfy assumptions?
model <- glmmTMB(formula = Adults ~ Trt + (1|Site/Trt), data = Data, family = "nbinom1")
summary(model)
plot(simulateResiduals(model))

model <- glmmTMB(formula = Adults ~ Trt + (1|Site/Trt), data = Data, family = "nbinom2")
summary(model)
plot(simulateResiduals(model))

model <- glmmTMB(formula = Adults ~ Trt + (1|Site/Trt), data = Data, family = "poisson")
summary(model)
plot(simulateResiduals(model))

model <- glmmTMB(formula = Adults ~ Trt + (1|Site/Trt), data = Data, family = "gaussian") # = "normal"
summary(model)
plot(simulateResiduals(model))

model <- glmmTMB(formula = log(Adults+1) ~ Trt + (1|Site/Trt), data = Data, family = "gaussian") # = "normal"
summary(model)
plot(simulateResiduals(model))


#options:
#1. ignore qqplot and use our glmm, it's good enough. Large number of observtions makes the test too sensitive.
#     i) the one site with very low counts and many zeroes is probably the problem
#2. just use the simple summary version with averaging to remove pseudoreplication
#3. explore zero-inflated models and other fancy things to find a way


#I could not find a solution, suggest option #2

```




