---
title: 'W2: data'
output:
  html_document:
    df_print: paged
---

First: compare R script vs. notebook

Goals:
-know ways to load in data just how you want it
-know how to diagnose and fix common problems
-basic cleanup to prep for graph or analysis
-best practices for datafiles (maybe a topic for later)

Philosophy:
-have ONE copy of your dataset. No v1, v2, _final, _finalfinal, _therealfinal, etc...
-use R to make calculations, transformations, subsets, summaries
-from now on I plan to share datafiles (increasingly required by journals!) and reproducible code for each published project
-science should be reproducible and open to scrutiny

```{r Chunk 1 - loading your data}

#use "setwd()" if your data file is not in the same directory as your R script

#LOADING with read.csv()
data_fig4 <- read.csv("data_week2_orpetEEAfig4.csv")
names(data_fig4) #notice the first cell has some weird characters, this the file encoding for csvs and we don't want to read it
data_fig4 <- read.csv("data_week2_orpetEEAfig4.csv", fileEncoding = 'UTF-8-BOM')


#INSTALLING and loading packages
library(readxl) #if you don't have this library, go to Tools, install packages


#LOADING with read_excel()
#simple way
data_fig4 <- read_excel("data_week2_orpetEEAfig4.xlsx")
#useful options
data_fig4 <- read_excel("data_week2_orpetEEAfig4.xlsx",
                   skip = 0, #skip rows if your dataset starts lower
                   guess_max = 10000, #prevents columns with many initial blanks importing as logical FALSE)
                   na = " ") # you can put a period in the quotes if that is how you code blanks.
#to load from a different sheet, use sheet = 

#GOOGLEDOCS
#library(googlesheets4)
#Data_scouting <- read_sheet("put the URL here", skip = 0)

```


```{r Chunk 2 - data cleanup and prep}

#(example) change weeks from a number to a character
data_fig4$Weeks <- as.character(data_fig4$Weeks)

#(example) re-order and change treatment names from alphabetical to order I will want on my figure legend
data_fig4$Elicitor <- factor(data_fig4$Elicitor,
                         levels=c("Harpin", "ASM"),       #I change the order of the levels (not the order of the data though)
                         labels=c("Harpin", "Actigard"))  #I change the name of ASM to Actigard

#(example) add a new column for a calculation (this is example is nonsense, but just to show you)
#philosophy: don't do calculation in excel! Keep your datafiles raw an unmanipulated
#adding two columns together
data_fig4$bogus_sum <- data_fig4$Foldchange + data_fig4$ddCT
#do a transformation
data_fig4$bogus_sqrt <- sqrt(data_fig4$bogus_sum + 100)
data_fig4$ddCT_negative <- -(data_fig4$ddCT) #I originally did this in excel, resulting in 0 instead of blank for one row

#(advanced) making a new column with "labels" for treatment I will want to use later in figure
#will explain "expression(paste())" later. The purpose is you can have a label with formatting like bold, superscripts etc...
data_fig4$Experimentlabel <- factor(data_fig4$Experiment, labels=c(
               "A"=expression(paste("27 April field spray")),
               "B"=expression(paste("23 May field spray")),
               "C"=expression(paste("Potted tree spray "))))

#subsetting
data_fig4_expA <- subset(data_fig4, Experiment == "A")
data_fig4_bogussubset <- subset(data_fig4, ddCT > 0 & Elicitor == "Actigard") #subset multiple conditions
data_fig4_bogussubset <- subset(data_fig4, Experiment == "A" | Experiment == "B") # use "|" for "or"


```

```{r Chunk 3 - a graph}

#makes use of:
##legend is in the order we asked for
##the labels we made earlier for panels
##could have but not really - weeks treated as a categroy instead of a continuous number, transformation to -ddCT

library(ggplot2)

ggplot(data = data_fig4, aes(x = Weeks, y = minusddCT)) +
  #geom_errorbar(aes(ymin = Foldchange-SEM, ymax = Foldchange+SEM, group = Elicitor), width = 0, position=position_dodge(width=0.9)) +
  #geom_point(stat = "identity", shape = 21, color = 'black', aes(fill = Elicitor), position=position_dodge(width=0.5)) +
  geom_hline(yintercept = 0, linetype = 'dashed', size = 0.25) +
  geom_point(shape = 21, size = 2, aes(group = Elicitor), position=position_dodge(width=0.6)) +
  geom_errorbar(stat = "summary", fun.data = mean_se, aes(group = Elicitor), position=position_dodge(width=0.6), width=0) +
  geom_point(shape = 23, size = 2, stat = "summary", fun = "mean", color = 'black', aes(fill = Elicitor), position=position_dodge(width=0.6)) +
  facet_wrap(~Experimentlabel, ncol = 1, scales = 'free',labeller = label_parsed) +
  scale_fill_manual(values=c('black', 'grey')) +
  geom_blank(data = subset(data_fig4, Experiment  == "A"), aes(y = 3)) +
  geom_blank(data = subset(data_fig4, Experiment  == "B"), aes(y = 4, x = '1')) +
  geom_blank(data = subset(data_fig4, Experiment  == "C"), aes(x = '7')) +
  scale_y_continuous(expression(paste("Relative ",italic("PR-1 "), "expression (-",paste(Delta),paste(Delta),"C"[T],")"))) +
  scale_x_discrete("\nWeeks post-treatment",breaks = c("0 (pre-spray)","1","3","7")) +
  theme_classic() +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 0.75) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 0.75) +
  guides(fill=guide_legend(ncol=1)) +
  theme(
        #strip.text = element_blank(),
        #strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        #panel.border = element_rect(colour = "black", fill = NA),
        axis.line = element_line(colour = 'black', size = 0.5),
        axis.ticks = element_line(colour = "black"),
        axis.text.x = element_text(color = 'black'),
        axis.text.y = element_text(color = 'black'),
        strip.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        legend.position=c(1,1), legend.justification=c(1,0),legend.box="vertical", legend.margin=margin(), legend.title = element_blank(), legend.key.size = unit(0.75, 'lines')
        #panel.spacing = unit(1.5, "lines")
        )

ggsave("Figure_4.tiff", width = 8, height = 12, units = "cm", dpi = 400, bg = "transparent")


```

best practices for datafiles (maybe a topic for later)
--avoid multiple tabs in an excel file




